Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2024-08-30 paper.tex
Remove non-existent images

To detect the direction of anisotropy we use an $N \times N$ (where $N$ is
dimensionality of the sample) covariance matrix
$\Sigma = \left\{ \sigma_{ij} \right\}$ for coordinates of points in the phase of
interest:
\begin{equation}
  \sigma_{ij} = \sum_{x \in \left\{
    \begin{array}{l}
      \text{Coordinates of points} \\
      \text{in the phase of interset}
    \end{array}
    \right\}} (x_i - \overline{x_i})(x_j - \overline{x_j})
\end{equation}
and decompose $\Sigma$ as $T \Lambda T^{-1}$ where $T$ is a unitary matrix and
$\Lambda$ is diagonal. $T$ can always be made so that it defines a rotation (by
multiplication of one of the columns by $-1$ if needed, so that
$\det T = 1$). Application of this rotation to the sample (or its inverse to the
pattern) makes the direction of anisotropy coaligned with the pattern, so
anisotropy can be better represented in the correlation function. For the sake
of visual clarity in this study we choose to rotate the image instead of the
pattern (\cref{fig:aniso}). Any rotation can potentially reduce accuracy of
computed functions due to necessity to interpolate the image during
rotation, and this important issue is explored later.
\begin{figure*}[ht]
  \centering
  \includegraphics[width=0.8\linewidth]{images/fig_rot.png}
  \caption{Examples of directional analysis of images with the help of $S_3$
    function when dealing with highly anisotropic media. The right-triangular
    pattern (\cref{fig:pattern}, right side) which we use in our algorithm is
    most sensitive to anisotropy which is directed along one of the
    axes. \\ From top down, left to right: an original image of an anisotropic
    pore; a rotated image of the pore, so it is positioned along the major axis;
    $S_3$ function for the original pore; $S_3$ function for the rotated pore
    with a vector $(0.92, -0.4)$ being a direction of elongation of the pore;
    $S_3$ function averaged across rotations similar to previously known
    algorithms \cite{berryman1988,SMITH1988176,malmir2018}.}
  \label{fig:aniso}
\end{figure*}

@@

The results of the comparison are shown in \cref{fig:comparison_patterns}. For
isotropic samples curves for directional and ensemble averaged, as well as for
right-angle isosceles triangle, basically collapse to each other with minimal
differences.
\begin{figure*}[ht]
  \centering \includegraphics[width=0.8\linewidth]{images/fig_compar.png}
  \caption[]{A comparison of computations of $S_3$ correlation function for the
    Sandstone sample using different patterns with and without random
    rotations. Left: evaluated with isosceles triangle pattern with the side $r$
    and the angle between edges $\phi$; sampling triangles are aligned with the
    plane XY.  Middle: evaluated with equilateral triangle pattern with the side
    $r$; sampling triangles are aligned with the plane XY.  Right: evaluated
    with equilateral triangle pattern with the side $r$. The result is averaged
    over 20 random rotations of the sampling triangle. Lower image: comparison
    of $S_3$ functions calculated with right triangle pattern with length of the
    both catheti $r$ and equilateral triangle pattern (both averaged over
    rotations and aligned with the plane XY).}
  \label{fig:comparison_patterns}
\end{figure*}

>>>>

To detect the direction of anisotropy we use an $N \times N$ (where $N$ is
dimensionality of the sample) covariance matrix
$\Sigma = \left\{ \sigma_{ij} \right\}$ for coordinates of points in the phase of
interest:
\begin{equation}
  \sigma_{ij} = \sum_{x \in \left\{
    \begin{array}{l}
      \text{Coordinates of points} \\
      \text{in the phase of interset}
    \end{array}
    \right\}} (x_i - \overline{x_i})(x_j - \overline{x_j})
\end{equation}
and decompose $\Sigma$ as $T \Lambda T^{-1}$ where $T$ is a unitary matrix and
$\Lambda$ is diagonal. $T$ can always be made so that it defines a rotation (by
multiplication of one of the columns by $-1$ if needed, so that
$\det T = 1$). Application of this rotation to the sample (or its inverse to the
pattern) makes the direction of anisotropy coaligned with the pattern, so
anisotropy can be better represented in the correlation function. For the sake
of visual clarity in this study we choose to rotate the image instead of the
pattern (\cref{fig:aniso}). Any rotation can potentially reduce accuracy of
computed functions due to necessity to interpolate the image during
rotation, and this important issue is explored later.
%\begin{figure*}[ht]
%  \centering
%  \includegraphics[width=0.8\linewidth]{images/fig_rot.png}
%  \caption{Examples of directional analysis of images with the help of $S_3$
%    function when dealing with highly anisotropic media. The right-triangular
%    pattern (\cref{fig:pattern}, right side) which we use in our algorithm is
%    most sensitive to anisotropy which is directed along one of the
%    axes. \\ From top down, left to right: an original image of an anisotropic
%    pore; a rotated image of the pore, so it is positioned along the major axis;
%    $S_3$ function for the original pore; $S_3$ function for the rotated pore
%    with a vector $(0.92, -0.4)$ being a direction of elongation of the pore;
%    $S_3$ function averaged across rotations similar to previously known
%    algorithms \cite{berryman1988,SMITH1988176,malmir2018}.}
%  \label{fig:aniso}
%\end{figure*}

@@

The results of the comparison are shown in \cref{fig:comparison_patterns}. For
isotropic samples curves for directional and ensemble averaged, as well as for
right-angle isosceles triangle, basically collapse to each other with minimal
differences.
%\begin{figure*}[ht]
%  \centering \includegraphics[width=0.8\linewidth]{images/fig_compar.png}
%  \caption[]{A comparison of computations of $S_3$ correlation function for the
%    Sandstone sample using different patterns with and without random
%    rotations. Left: evaluated with isosceles triangle pattern with the side $r$
%    and the angle between edges $\phi$; sampling triangles are aligned with the
%    plane XY.  Middle: evaluated with equilateral triangle pattern with the side
%    $r$; sampling triangles are aligned with the plane XY.  Right: evaluated
%    with equilateral triangle pattern with the side $r$. The result is averaged
%    over 20 random rotations of the sampling triangle. Lower image: comparison
%    of $S_3$ functions calculated with right triangle pattern with length of the
%    both catheti $r$ and equilateral triangle pattern (both averaged over
%    rotations and aligned with the plane XY).}
%  \label{fig:comparison_patterns}
%\end{figure*}
